---
title: "week7_lab"
author: "Satie Airame"
date: "11/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objects and classes

# Vectors

### Creating a vector

```{r}
# Numeric vector
my_vector <- vector(mode = "numeric", length = 100) # creates only zeros unless otherwise specified, length can be anything you want

# Logical vector
my_second_vector <- vector(mode = "logical", length = 20) # contains pipe logical (TRUE/FALSE), default is FALSE (0), also known as Boolean operators
```

### Vector properties

#### Length
```{r}
length(my_vector) # call the length of the vector
length(my_second_vector)
```

#### Class

```{r}
# What is is my vector?
class(my_vector)
class(my_second_vector)
```
#### Indexing (accessing) vectors

```{r}
#Using c to create a vector

my_vector <- c(1:50)

# What's the first element?
my_vector[1]

# What are all of the elements except the first one?
my_vector[-1]

# What's the seventh element?
my_vector[7]

# Select a range from a vector (from the third object to the seventh object)
my_vector[3:7]
```

## Matrices (vectors with length in 2 directions)

### Creating matrices

```{r}
# Create a 10 x 10 matrix full of NA's
my_matrix <- matrix(data = 1:100, nrow = 20, ncol = 5) # Store numbers from 1-100 in 20 rows in 5 columns

```
## Matrix properties

#### Dimensions (dim)

```{r}
dim(my_matrix) # Gives 2 numbers: first is number of ROWS, second is number of COLUMNS
```

#### Length

```{r}
# How many elements are there?
length(my_matrix) # Shows number of elements
```

#### Class

```{r}
class(my_matrix)
```

# An array is a bidimensional matrix

#### Indexing

```{r}
my_matrix[2,] # Asks what values are in the second ROW, the first number always indicates ROWS. Second space (COLUMNS) is empty, so R assumes you want all of the values in row 2.

# Index a column
my_matrix[,3]# Asks what values are in the third COLUMN, the second number always indicates COLUMNS. First space (ROWS) is empty, so R assumes you want all of the values in COLUMN 3.

# Index a given cell
my_matrix[8,3] # Index value for Row 8 and column 3

# Index and replace
my_matrix[8,3] <- 100000 # Put 100000 in row 8 and column 3

# Replace an entire column
my_matrix[,5] <- 10 # Replace everything in column 5 with "10"


```

# Control Flow

## 'for' loops # May you do something until condition is met

variable(counter), sequence, expression

```{r}
for (i in 1:10) # specify variable i and specify sequence i = 1, i = 2, i = 3, etc.
  print("Hello world!")
```

# Generating random numbers in R

```{r}
rnorm(n=5, mean = 10, sd = 2) #random normal distribution, 5 observations, with mean 10, sd = 2
```

# Let's combine all of the above.

We will use "numeric" to create an empty vector.
We will use "for" to iterate along items of that empty vector
We will use "[ ] to access each slot and fill-in thta empty vectore
We will use "rnorm" to sample values from a random normal distribution
We will plot the trend

Discrete Time Logistic Growth Equation

$$N_{t}=N_{t-1} + r\left(1-\frac{N_{t-1}}{K}\right)$$

where r is the intrinsic growth rate
and K = carrying capacity 

Density dependence due to resource limitation.

$$
r\sim norm(\mu = 1.5, \sigma = 0.8)
$$

## Writing the code down

```{r}
# Set up model parameters
K <- 100 #individuals
nsteps <- 100 #years
mu <- 1.4
sigma <- 1

# Set up model objects

N <- vector(mode = "numeric", length = nsteps)
N[1] <- 25

# Model it
for(t in 2:nsteps){ # create repeating function beginning at year 2
  r <- rnorm(n = 1, mean = mu, sd = sigma) # create random value of r
  N[t] <- N[t-1] + r * (1 - (N[t-1]/K))  # t is counter from for loop
}
  
# Visualize it
plot(N)

```
# Something more
We will use "matrix" to create an empty matrix (rows are time, columns are runs)
We will repeat the process

```{r}
# Set up model parameters
K <- 100 #individuals
nsteps <- 100 #years
mu <- 0.4
sigma <- 1

# Set up model objects

N <- matrix(data=NA, nrow = nsteps, ncol = 100) # NA is better than 0 because NA means no value where 0 means there were no individuals. 
N[1,] <- 25 # need to add a comma to specify this is a matrix

# Model it
for(t in 2:nsteps){ # create repeating function beginning at year 2
  r <- rnorm(n = 100, mean = mu, sd = sigma) # create random value of r
  N[t,] <- N[t-1,] + r * (1 - (N[t-1,]/K))  # t is counter from for loop, need to add commas to indicate matrix 
}
  
# Visualize it
matplot(N, type = "l") # indicate that type is linear

```

# Bring in your friends

Covert it to things we know about: ggplot

```{r}
# Load tidyverse
 library(tidyverse)
#Create data.frame
my_data <- as_tibble(N) %>% 
  mutate(time = 1:nsteps) %>% 
  gather(run, N, -time)
# Visualize it
ggplot(data = my_data, mapping = aes(x = time, y = N))+
  geom_line(aes(group = run), size = 0.1, alpha = 0.5)+ # one geom_line object for each model run
  stat_summary(geom = "line", fun = mean, color = "red", size = 1)+ # create average line of all runs, use the function ("fun") mean and color red
  theme_bw()+
  labs(x = "Time (years)", y = "Population size (N)")
 

```







